---
title: "Getting the data on small scale"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting the data on small scale}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r warning=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(httr)
library(readr)
library(dplyr)
library(tidyr)
library(purrr)
library(usethis)
library(ggplot2)

devtools::load_all()
```

# EXTRACT DATA

## Trams GPS coordinates
Data was taken from API, with calls every 60 seconds (refresh rate of API), over 3 hours. Data limited to line 7. Data access date 07-02-2022

```r
get_trams_gps(
  line_number = 7, 
  time_interval = 60, 
  max_calls = 181, 
  output = "../data-raw/tram-7-3h.tsv"
  )
```

### CLEAN TRAMS GPS DATA
```{r echo=TRUE}
trams <- read_delim(
  file = "../data-raw/tram-7-xxx-int15s.tsv",
  skip = 1, 
  col_names = FALSE,
  show_col_types = FALSE)

# remove row indexes
trams <- trams[2:ncol(trams)]

colnames(trams) <- c("line", "lon_tram", "vehicle_number", "time", "lat_tram", "brigade", "get_timestamp")

head(trams)
```


## Tram and bus stops coordinates

Next step is to identify where the stops are. After api call we receive ugly nested list that I stored in JSON format.
```{r eval=FALSE, include=FALSE}
url <- paste0(
  "https://api.um.warszawa.pl/api/action/dbstore_get?id=1c08a38c-ae09-46d2-8926-4f9d25cb0630",
  "&apikey=", Sys.getenv("UM_WAW_PERSONAL_API_KEY")
  )

przystanki <- GET(url)

przystanki <- content(przystanki)

# jsonlite::write_json(przystanki, "przystanki.json", pretty = TRUE)
```

That looks like this:
```json
{
  "result": [
    {
      "values": [
        {
          "value": ["1001"],
          "key": ["zespol"]
        },
        {
          "value": ["01"],
          "key": ["slupek"]
        },
        {
          "value": ["Kijowska"],
          "key": ["nazwa_zespolu"]
        },
        {
          "value": ["2201"],
          "key": ["id_ulicy"]
        },
        {
          "value": ["52.248455"],
          "key": ["szer_geo"]
        },
        {
          "value": ["21.044827"],
          "key": ["dlug_geo"]
        },
        {
          "value": ["al.Zieleniecka"],
          "key": ["kierunek"]
        },
        {
          "value": ["2021-10-08 00:00:00.0"],
          "key": ["obowiazuje_od"]
        }
      ]
    },
...
```

### CLEAN STOPS DATA

After the clean-up of data into the tabular format
```{r}
przystanki_df <- get_all_stops_gps()
przystanki_df |> head()
```





## GET TIMETABLES

In order to get timetables, we need to know which tram/bus stops at which stop to generate the correct API call.

#### DO NOT RE-RUN THIS CELL UNLESS YOU WANT TO UPDATE DATA
```{r eval=FALSE, include=FALSE}

unique_stop_instances <- przystanki_df |> 
  select(team, pole) |> 
  distinct()

busstops_lines <- get_all_lines_at_stop(unique_stop_instances)

# last save 25-03-2022
write_rds(busstops_lines, "bustops_lines.rds", compress = "xz")
```

Get only stops where line 7 stops
```{r}
busstops_lines <- read_rds("bustops_lines.rds")
busstops_lines <- select(busstops_lines, team, pole, data)


stops_7 <- busstops_lines |> 
  #filter(team == "1001", pole == "01") |> 
  mutate(lines = map(data, ~ .x$value)) |> 
  unnest_longer(lines, values_to = "line") |> 
  filter(line == "7") |> 
  select(-data)
head(stops_7)
```
Keep stops details only for line 7
```{r}
stops_details_7 <- semi_join(przystanki_df, stops_7, by = c("team", "pole"))
stops_details_7 <- stops_details_7 |> 
  rename(lon_stop = lon,
         lat_stop  = lat)
head(stops_details_7)
```
Now we can get timetables for all stops of line 7
```{r get timetables}
line <- 7

unique_stop_instances_7 <- stops_details_7 |>  
  select(team, pole) |> 
  distinct()

timetables_7 <- get_timetables(unique_stop_instances_7, line)

readr::write_csv(timetables_7, "timetables_7_sunday_10-04-2022.csv")
```


Identify pętle i zajezdnie
```{r}

stops_details_7 |> pull(direction) |> unique() |> sort()

# zajezdnie ZET

# pętle dla 7 manualnie "P+R Al.Krakowska" i "Kawęczyńska-Bazylika"

```

# VISUALIZE
```{r}
# sprawdzenie przystankóœ na pl narutowicza
stops_details_7 |> 
  filter(team_name == "pl.Narutowicza") |> 
  ggplot(aes(x = lon_stop, y = lat_stop)) +
  geom_point()+
  geom_text(aes(label = pole))

```
```{r }
# sprawdzam czy trasa zawsze zgadza się z założeniami ze następny przystanek można brać z "direction"
timetables_7 %>% 
  mutate(filtered = map(timetables_7$timetable, ~ filter(., trasa == "TZ-KAW3"))) %>%
  unnest(cols = filtered) |>
  mutate(czas = lubridate::hms(czas)) |> 
  arrange(czas) |> 
  select(team, pole, kierunek, czas) |> 
  left_join(select(stops_details_7, team, pole, team_name, direction), by = c("team", "pole"))
```

Wniosek - __NIE__. Trzeba zebrać wszystkie trasy posortować po czasie przyjazdu i ustalić kolejność przystanków na ich zasadzie. Np dla trasy TZ-KAW3 do zajezdni mokotów direction jako następny przystanek nie sprawdza się, gdy tramwaj odbija do zajezdni.

Nie udało się opracować jednego sposobu na określenie kolejności przystanków per trasa bo brakuje zmiennej, która oddzielałaby przejazdy tej samej brygady po tej samej trasie w ciągu dnia. 

Kolejność przystanków określona na podstawie timetable dla rozkładu weenekdowego (czy to ma znaczenie?) 
```{r wytyczenie kolejności przystankóœ na trasie}
feats <- c("trasa","czas_rozkladu", "brygada", "team", "pole",  "kierunek")

clean_timetable_7 <- timetables_7 |> 
  unnest(cols = timetable) |> 
  filter(!is.na(czas)) |> # remove rows for which api failed to get data
  mutate(
    czas_rozkladu = lubridate::hms(czas),
    brygada = as.numeric(brygada)) |> 
  select(feats) |> 
  arrange(trasa, czas_rozkladu, brygada, team, pole) |> 
  left_join(select(stops_details_7, team, pole, team_name), by = c("team", "pole"))  |> 
  group_by(trasa, brygada) |> 
  arrange(trasa, brygada, czas_rozkladu) 

clean_timetable_7 |> pull(trasa) |> unique()

# TD-2OKE - występuje raz z zajezdni praga na okęcie.
td_2OKE_order <- clean_timetable_7 |> 
  ungroup() |> 
  filter(trasa == "TD-2OKE") |> 
  transmute(
    trasa = trasa,
    team = team,
    pole = pole,
    team_name,
    stop_order = row_number())

# TD-3OKE - występuje kilka razy dlatego zlimitowane do brygady
# z zajezdni mokotów na okęcie
td_3OKE_order <- clean_timetable_7 |> 
  ungroup() |> 
  filter(trasa == "TD-3OKE", brygada ==  1) |> 
  transmute(
    trasa = trasa,
    team = team,
    pole = pole,
    team_name,
    stop_order = row_number())

# TP-KAW - główna trasa okęcie --> kawęczyńska 
# występuje wiele razy dlatego zlimitowane do brygady i ograniczone czasowo
# żeby wyłapać jeden przejazd

tp_kaw_order <- clean_timetable_7 |> 
  ungroup() |> 
  filter(
    trasa == "TP-KAW", 
    brygada == 1, 
    czas_rozkladu < lubridate::hm("7:00")
    ) |> 
  transmute(
    trasa = trasa,
    team = team,
    pole = pole,
    team_name,
    stop_order = row_number()
    )

# TP-OKE - główna trasa kawęczyńska --> okęcie 
tp_oke_order <- clean_timetable_7 |> 
  ungroup() |> 
  filter(
    trasa == "TP-OKE", 
    brygada == 1,
    czas_rozkladu < lubridate::hm("8:00")
    ) |> 
  transmute(
    trasa = trasa,
    team = team,
    pole = pole,
    team_name,
    stop_order = row_number()
    )

# TZ-OKE2 - trasa z okęcia do zajezdni ZET-PRAGA - jedna brygada
tz_oke2_order <- clean_timetable_7 |> 
  ungroup() |> 
  filter(
    trasa == "TZ-OKE2"
    ) |> 
  transmute(
    trasa = trasa,
    team = team,
    pole = pole,
    team_name,
    stop_order = row_number()
    )

# TZ-OKE3 - trasa z okęcia do zajezdni ZET-MOKOTÓW - dwie brygady
tz_oke3_order <- clean_timetable_7 |> 
  ungroup() |> 
  filter(
    trasa == "TZ-OKE3",
    brygada == 1
    ) |> 
  transmute(
    trasa = trasa,
    team = team,
    pole = pole,
    team_name,
    stop_order = row_number()
    )


trasy <- bind_rows(
  td_2OKE_order,
  td_3OKE_order,
  tp_kaw_order,
  tp_oke_order,
  tz_oke2_order,
  tz_oke3_order
)

readr::write_csv(trasy, "trasy.csv")

head(trasy)
```

```{r}
head(trams)
```
```{r}
head(clean_timetable_7)
```
```{r where is my tram - scratches}
x <- trams |> 
  mutate(
    czas = as.character(time) |> 
      stringr::str_extract(pattern = "\\d\\d:\\d\\d:\\d\\d$") |> 
      lubridate::hms()
  ) |> 
  #select(get_timestamp) |> 
  filter(as.character(get_timestamp) == "2022-03-23 20:16:07",
         brigade == 1) 
  #left_join(clean_timetable_7, by = c("brigade"="brygada"))
  #left_join(stops_details_7, by = c("brigade"="brygada"))


head(stops_details_7)
# find closest entry in timetable ==> znaleźć obecną trasę
obecna_trasa <- x |> 
  left_join(clean_timetable_7, by = c("brigade"="brygada")) |> 
  mutate(
    czas_diff = abs(as.difftime(czas - czas_rozkladu))
  ) |> 
  filter(czas_diff == min(czas_diff)) |> 
  slice_head(n=1) |> 
  pull(trasa)

przystanki_obecna_trasa <- trasy |> filter(trasa == obecna_trasa)
stop_det_current <- semi_join(
  stops_details_7, 
  przystanki_obecna_trasa,
  by = c("team", "pole"))
# find closest stop?
dist_to_stop <- x |> 
  left_join(stop_det_current, by = character()) |> 
  rowwise() |> 
  mutate(
    dist_to_stop = geosphere::distHaversine(
      c(lon_tram, lat_tram), 
      c(lon_stop, lat_stop)
      )
  ) 
```


### Visualise data

```{r trasy, fig.width=10, fig.height=10}
trasy <- readr::read_csv("trasy.csv")
trasy |> 
  left_join(select(stops_details_7, team, pole, lat, lon), by = c("team", "pole")) |> 
  ggplot(aes(x = lon_stop, y = lat_stop, color = trasa)) + 
  geom_point(color = "black") +
  geom_path(show.legend = FALSE)+
  ggrepel::geom_label_repel(aes(label = team_name), size = 3, show.legend = FALSE)+
  #geom_text(aes(label = team_name), size = 2)+
  facet_wrap(~ trasa)

```

```{r plot line 7, fig.height=10, fig.width=10}
tracked_tram <- trams |> 
  filter(as.character(get_timestamp) == "2022-03-23 20:16:07" & brigade == 1)

ggplot(data = trams, aes(x = lon_tram, y = lat_tram)) +
  geom_point(alpha = .1, show.legend = FALSE) +
  geom_point(
    data = tracked_tram,
    aes(x = lon_tram, y = lat_tram),
    color = "red", 
    size = 3) +
  geom_point(data = stops_details_7, aes(x = lon_stop, y = lat_stop), color = "orange") +
  ggrepel::geom_text_repel(
    data = stops_details_7, 
    aes(x = lon_stop, y = lat_stop, label = paste(team_name, pole)),
    size = 3, 
    max.overlaps = 20) +
  ggtitle("Trams and tram stops location")

```

### Plot stops
```{r}
przystanki_df |>
  ggplot(aes(x = lon, y = lat)) +
  geom_point(alpha = .1, show.legend = FALSE)+
  ggtitle("All bus and tram stops in Warsaw")
```

### Distance between two coordinates in straight line in meters
```{r}
geosphere::distHaversine(c(21.04766, 52.25201), c(21.04949, 52.25262))
```

### How to detect if tram is close to / at the stop?

1. Assume that tram stop is a circle of a given radius
2. If distance to the stop is =< to the radius then assume stop

Problem: what if tram is caught after the stop? 
Solution: 
1. distance to the next stop (must know which and where is next stop) is shorter than previously?
2. di

